Case-based Arabic Morphological Analysis

### 2010-12-16 trunk/tex/pr.tex
final

This report describes the progress of the work carried out in the LNCSR funded project entitled: {\bf ``Relational Queries for Arabic Text Mining"}. 
\\

This document is organized as follows:
\begin{enumerate}
\item Summary of work conducted during the first year of the project
\item Goals for next year
\item Budget for year 2011
\item Submission to: ``ACL/HTL 2011"
\end{enumerate}
\pagebreak

@@

At the American University of Beirut (AUB), we
started work on an Arabic text mining framework~\cite{ATMine09}
that is available now online. The source for this framework comprises code of the order of tens of thousands lines.
We have successfully developed ``SARF", a case-based morphological 
analyzer, which constitutes capabilities of both Lexical and Syntax Analysis. We have been able to verify that both in terms of speed and accuracy our analyzer outperformed 2 state-of-art analyzers: Buckwalter ~\cite{Buckwalter:02} and ElixirFM~\cite{Otakar:07}. Supporting statistics are present in the attached paper described in Section~\ref{s:paper}. In addition, we have applied SARF on a literature hadith case study.

@@

It is clear that our first year of research on this topic has been fruitful. For next year, we expect to complete the hadith literature case study and to start and complete the security case study. We will package the components built during the development of these case studies as independent text mining entity and relational extractors to be included in the framework to be completed. Moreover, we expect to be able to enhance the morphological analyzer in terms of speed and accuracy. We will also investigate methods to extend the internal structures of our analyzer to be able to represent a formal linguistic
computational model of the Arabic language. In what follows we describe our expectations in terms of each case-study.

@@

The diagram in Figure~\ref{f:security} describes our 
high level vision for what the solution of the 
security case study
will look like in the framework. 
The inputs are two separate sets of documents. 
The first set includes crime scene investigation reports, 
and the second set includes a 
interrogation reports with crime suspects as well as 
a database of previous interrogation reports. 
The system also takes as input a database of geographical
locations and a database of personal records. 
The aim is to check for inconsistencies in the interrogation
reports and to identify possible connections between the suspects
and the crime scene.

@@

In the table below we summarize our itemized budget for next year:

@@

 \hline
 One part-time research assistant for 8 month: LL. 750K/month  & LL. 6000K  \\
 Arabic Corpus Lisence & LL. 1500K \\
Miscellaneous  & LL. 500K \\
\hline
 Total  & LL. 8000K \\
 \hline
\end{tabular} 
\end{table}

@@

We have submitted the attached paper to the “The 49th Annual Meeting of the Association for Computational Linguistics: Human Language Technologies (ACL/HLT 2011)”. The paper is entitled: “Case-based Arabic Morphological Analysis”.

>>>>

In this report we discuss our progress and work on the 
{\bf Relational Queries for Arabic Text Mining} project 
funded by LNCSR.
\\

This document is organized as follows:
\begin{enumerate}
\item Summary of work conducted during the first year of the project
\item Goals for next year
\item Budget for year 2011
\item Attached submission to ``ACL/HTL 2011"
\end{enumerate}
\pagebreak

@@

At the American University of Beirut (AUB), we
started work on an Arabic text mining framework~\cite{ATMine09}
that is available now online. 
ATMine is around 20K lines of C++, perl, and sql code and uses 
middleware from the open source community such as MySQL, Qt, and 
double array trie.

Arabic morphological analyzers~\cite{Sughaiyer:04}
consider an Arabic word and its internal structure composed of 
several {\em morphemes}. 
A morpheme is a {\em stem} or an {\em affix}.
An affix is a {\em prefix, suffix,} or an {\em infix}.
The analysis of one word may lead to several possible
solutions.
\vocalize
For instance, the word 
\RL{'a.hmadH}~\footnote{In this document, we use the default 
ArabTeX transliteration style ZDMG.}
may have two valid morphological analyses. 
The word means ``I praise him'' when
the letter \RL{'a} is a prefix and  it means
``his Ahmad'' when 
the letter \RL{'a} is part of the proper noun stem 
\RL{'a.hmad}.

Current morphological analyzers such as 
Buckwalter~\cite{Buckwalter:02},
Beesely~\cite{Beesley:01}, SAMA~\cite{Kulick:10},
and ElixirFM~\cite{Otakar:07} exist.
These analyzers are used in several open source spell checkers as 
well as NLP frameworks~\cite{Col09}.
They take as input white space delimited tokens~\cite{Kulick:10},
consider them as words,
and enumerate all possible solutions. 
This approach has several problems. 
First, a white space delimited token may have 
more than one word.
Second, the exhaustive enumeration may hurt performance and may
not be necessary or appropriate
in some case studies~\cite{Maamouri:10}. 
Other morphological analyzers such as 
Amira~\cite{Diab:07,Benajiba:07},
MAGEAD~\cite{Habash:05}, and MADA+TOKAN~\cite{Habash:09} 
use machine learning and support vector machines (SVM) 
to enhance the accuracy at the expense of efficiency.

We have successfully developed ``SARF", a case-based morphological 
analyzer with lexical and syntax analysis capabilities. 
SARF outperforms in terms of efficiency and accuracy 
the Buckwalter~\cite{Tim04} and the ElixirFM~\cite{Otakar:07} 
analyzers. 
We report on the detailed results in the attached manuscript.
We discuss the attached manuscript in Section~\ref{s:paper}. 
In addition, we have applied SARF on a hadith literature case study.

@@

Our first year of research on 
Arabic text mining proved to be 
fruitful. 
For the next year, we expect to complete the hadith literature case 
study and to start and complete the security case study. 
We will package the components built during the development of these 
case studies as independent text mining entity and relational 
extractors to be included in the framework.
Moreover, we expect to be able to enhance the morphological analyzer in 
terms of speed, accuracy, and semantic capabilities. 
We will also investigate methods to extend the internal structures 
of our analyzer to be able to represent a formal linguistic
computational model of the Arabic language. 
In what follows we describe our expectations in terms of each case-study.

@@

The diagram in Figure~\ref{f:security} describes our 
high level vision for what the solution of the 
security case study
will look like in the framework. 
The inputs are two separate sets of documents. 
The first set includes crime scene investigation reports, 
and the second set includes a 
interrogation reports with crime suspects as well as 
a database of previous interrogation reports. 
We will first try our system on dummy data with seeded relations.
Once the system is ready, we will deliver it to local police
and intelligence agencies to try it out on real digitized 
reports.

The system also takes as input a database of geographical
locations and a database of personal records. 
The aim is to check for inconsistencies in the interrogation
reports and to identify possible connections between the suspects
and the crime scene.

@@

In the table below we summarize our itemized budget for year 2011.

@@

 \hline
 One part-time research assistant for 8 month: LL. 750K/month  & LL. 6000K  \\
 Arabic Corpus License & LL. 1500K \\
Miscellaneous  & LL. 500K \\
\hline
 Total  & LL. 8000K \\
 \hline
\end{tabular} 
\end{table}

@@

We have submitted the attached paper to the “The 49h Annual Meeting of the Association for Computational Linguistics: Human Language Technologies (ACL/HLT 2011)”. The paper is entitled: “Case-based Arabic Morphological Analysis”.
